---
title: '**DSA/ISE-5103**'
subtitle: '**Guidelines for Assignment Reports**' 
author: 'By Alexander Rodriguez'
date: 'Due to August 31, 2017'
output: 
  pdf_document:
    highlight: haddock
documentclass: article
geometry: margin = 0.8in
urlcolor: blue
---

# R Markdown

* R Markdown is a format for writing reproducible, dynamic reports with R. Use it to embed R code and results into slideshows, pdfs, html documents, Word files and more (i.e. a cool way to present your reports). For more about it, use this [link](http://rmarkdown.rstudio.com/).
* Headers in Markdown start #, and code starts with ```. For a summary of the most important commands of R Markdown, search on internet for **R Markdown Cheat Sheet**.
* Compile R Markdown files with the Knit button right above the code editor.

# About assignments for DSA/ISE 5103

* Make it look professional. Be suscint and straightforward. Only show code and results that are valuable to answer the questions or drive your conclusions.
* You are required to use R Markdown for your first assignment. You can opt to use other tool for the rest of the assignments.
* All assignment submissions need to be in **PDF format** and you have to **attach your R code**.
* Your submitted code has to have the following structure for each problem/question: visible problem number, all the libraries used for that , properly commented code. Failing to follow this structure will result in grade deduction.
* Assignment reports have a **limit of pages**. If you exceed them, some points will be deducted.

# About this document

This document is for guiding you how to hand in your homeworks with R Markdown. Some important remarks:

* You will find some useful commands especially in the code chucks. **Read the comments in the code chunks!** They are in the Rmd file that created this PDF.
* This version is compilable only for PDF, but the commands in the code chunks work also for Word and HTML.
* If you want to directly create PDFs, you need to download TeX. You can follow this [link](https://www.latex-project.org/get/) for instructions. You can also create PDFs by converting from Word to PDF.
* Don't compile this template in a shared folder like OneDrive or Dropbox because it won't work.

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
# By setting echo to false, the r code in this chuck won't appear on your report
# By setting warning and message to false, the annoying messages won't appear

# All the libraries statement that your are using have to go in the top of the file
library(ggplot2) # for data visualization
library(dplyr) # for data manipulation
```

\pagebreak <!-- this is a LaTex command that works on Markdown -->

# Example 1: Using code chuncks the right way

### Embedding results in plain text

In many times, the answers for questions are straightforward. You can make the calculations in a chunk and show the results embeded in your plain text paragraphs. Use *include = FALSE* at the top of the code chunk.

```{r, include=FALSE}
# If we want to compute this without showing the code neither the results use include = FALSE
m <- mean(mtcars$hp) # calculation of mean
s <- sd(mtcars$hp) # calculation of standard deviation
```

Now that you calculated these values, just call it by name of the variable and it will be printed in your report once you compile it. Here an example:

"The computed mean for the attribute hp is `r round(m, 2)` and its standard deviation is `r round(s, 2)`. From this we can conclude that ..."

### Showing only results

If you want to show ONLY the results as R output, do the following (this will not show the R code to make them happen):

```{r, echo=FALSE}
# Setting echo to false will make your results appear (if you are showing them), but not the code
quantile_hp <- quantile(mtcars$hp) # taking quantile for hp
quantile_hp # showing the results from the previous calculation
```

### Showing only R code

If you want to show the R code because you think it's important, but not to evaluate it (i.e. nothing happens when running this code chunk), then use *eval = FALSE*:

```{r, eval=FALSE}
# Set eval to false to avoid evaluation the code chuck
scale_0to1 <- function(x){(x-min(x))/(max(x)-min(x))} # creating a function to scale from 0 to 1
```

### Showing R code along with results

If you consider that some R code and results are needed to be remarked (think twice before you do it), set echo to TRUE and show it:

```{r, echo=TRUE}
# By default your code and results will appear
glimpse(mtcars) # command similar to str 
```

\pagebreak <!-- this is a LaTex command that works on Markdown -->

# Example 2: Shrinking plots to make them fit

It's a common problem that your plots are going to be too big to look nice in your report. Like the following plot: 

```{r, echo=FALSE}
# By setting echo to false, the r code in this chuck won't appear on your report

# Plotting weight vs miles per galon
arc <- ggplot(mtcars, aes(x=wt, y= mpg, col = factor(cyl))) + 
  geom_point(size = 4.5, alpha=0.6) +
  geom_smooth(method = "lm", se = F) +
  xlab("Weigth (lb/1000)") + ylab("Miles/US gallon") + 
  labs(col= "Cylinders") +
  stat_smooth(method = "lm", se = F) + 
  stat_smooth(method="lm", se=F, col = "black", aes(group=1)) +
  theme_classic()
arc # call the object to show the plot
```

You can shrink the plot easily with the *fig* commands, and you'll have set the figure width and height.

```{r, fig.width=4, fig.height=3, echo=FALSE}
# You may change the width and the height for your convenience (trial and error)
arc
```

We can even make two plots fit in the same space (as two columns) by using the right width and height.

```{r, fig.width=3.5, fig.height=2.5, echo=FALSE}
# Just change the width and height
arc
arc
```

Now that you have the perfect plot in the suitable space, **do never forget to comment about it!** If you don't comment/use them for your analysis, then there is no purpose to keep them in your report.

# Final remarks

* Remind, your report DOES NOT have to contain unnecesary code. Save space for your analysis and results that help your analysis.
* Each piece of results or plot that you are presenting in the report should come along with some analysis.
* If you have several interesting results, use a table or something creative to show them. Search on internet table tools like [this](http://www.tablesgenerator.com/markdown_tables).
* Do not forget to submit your R code along with your report. If you used R Markdown, you are good to upload your .Rmd file instead of the .R file as long as it has all the code you used.

Thanks for reading. Find the most updated version in [Github](https://github.com/arodriguezca/data-analytics_course) and share it if was useful for you!
